Improving Software Maintainability through Automated Refactoring of Code Clones

### 2019-08-14 icse/paper.tex
contents

\subsection{Context Mapping}
After clones are detected, we map the context of these clones. We have identifier three properties of clones as their context: relation, location and contents. We identify categories for each of these properties, to get a detailed insight in the context of clones.

\subsubsection{Relation}
% The rationale for our categories regarding clone relations.
Clone instances in a clone class can have a relation with each other through inheritance. This relation has a big impact on how the clone should be refactored \cite{fontana2015duplicated}. We define the following categories to map the relation between clone instances in a clone class. These categories do not map external classes (classes outside the project, for instance belonging to a library) unless explicitly stated:
\begin{itemize}
    \item \textbf{Common Class}: All clone instances are in the same class.
    \begin{itemize}
        \item \textbf{Same Method}: All clone instances are in the same method.
        \item \textbf{Same Class}: All clone instances are in the same class.
    \end{itemize}
    \item \textbf{Common Hierarchy}: All clone instances are in the same inheritance hierarchy.
    \begin{itemize}
        \item \textbf{Superclass}: Clone instances reside in a class or its parent class.
        \item \textbf{Sibling Class}: All clone instances reside in classes with the same parent class.
        \item \textbf{Ancestor Class}: All clone instances reside in a class, or any of its recursive parents.
        \item \textbf{First Cousin}: All clone instances reside in classes with the same grandparent class.
        \item \textbf{Same Hierarchy}: All clone instances are part of the same inheritance hierarchy.
    \end{itemize}
    \item \textbf{Common Interface}: All clone instances are in classes that have the same interface.
    \begin{itemize}
        \item \textbf{Same Direct Interface}: All clone instances are in a class that have the same interface.
        \item \textbf{Same Class}: All clone instances are in an inheritance hierarchy that contains the same interface.
    \end{itemize}
    \item \textbf{Unrelated}: All clone instances are in classes that have the same interface.
    \begin{itemize}
        \item \textbf{No Direct Superclass}: All clone instances are in classes that have the Object class as parent.
        \item \textbf{No Indirect Superclass}: All clone instances are in a hierarchy that contains a class that has the Object class as parent.
        \item \textbf{External Superclass}: All clone instances are in classes the same external class as parent.
        \item \textbf{Indirect External Ancestor}: All clone instances are in a hierarchy that contains a class that has an external class as parent.
    \end{itemize}
\end{itemize}
Based on these relations, we determine where to place the cloned code when refactored. The code of clones that have a \textit{Common Class} relation can be refactored by placing the cloned code in this same class. The code of clones with a \textit{Common Hierarchy} relation can be placed in the intersecting class in the hierarchy (the class all clone instances have in common as an ancestor). The code of clones with a \textit{Common Interface} relation can be placed in the intersecting interface, but in the process has to become part of the classes' public contract. The code of clones that are \textit{Unrelated} can be placed in a newly created place: either a utility class, a new superclass abstraction or an interface.

\subsubsection{Contents}
% The rationale for our categories regarding clone contents.

\begin{enumerate}
  \item \textbf{Full Method/Constructor/Class/Interface/Enumeration:} A clone that spans a full class, method, constructor, interface or enumeration, including its declaration.
  \item \textbf{Partial Method/Constructor:} A clone that spans (a part of) the body of a method/constructor.
  \item \textbf{Several Methods:} A clone that spans over two or more methods, either fully or partially, but does not span anything but methods.
  \item \textbf{Only Fields:} A clone that spans only global variables.
  \item \textbf{Other:} Anything that does not match with above-stated categories.
\end{enumerate}

>>>>

\subsection{Context Mapping}
After clones are detected, we map the context of these clones. We have identifier three properties of clones as their context: relation and contents. We identify categories for each of these properties, to get a detailed insight in the context of clones.

\subsubsection{Relation}
% The rationale for our categories regarding clone relations.
Clone instances in a clone class can have a relation with each other through inheritance. This relation has a big impact on how the clone should be refactored \cite{fontana2015duplicated}. We define the following categories to map the relation between clone instances in a clone class. These categories do not map external classes (classes outside the project, for instance belonging to a library) unless explicitly stated:
\begin{itemize}
    \item \textbf{Common Class}: All clone instances are in the same class.
    \begin{itemize}
        \item \textbf{Same Method}: All clone instances are in the same method.
        \item \textbf{Same Class}: All clone instances are in the same class.
    \end{itemize}
    \item \textbf{Common Hierarchy}: All clone instances are in the same inheritance hierarchy.
    \begin{itemize}
        \item \textbf{Superclass}: Clone instances reside in a class or its parent class.
        \item \textbf{Sibling Class}: All clone instances reside in classes with the same parent class.
        \item \textbf{Ancestor Class}: All clone instances reside in a class, or any of its recursive parents.
        \item \textbf{First Cousin}: All clone instances reside in classes with the same grandparent class.
        \item \textbf{Same Hierarchy}: All clone instances are part of the same inheritance hierarchy.
    \end{itemize}
    \item \textbf{Common Interface}: All clone instances are in classes that have the same interface.
    \begin{itemize}
        \item \textbf{Same Direct Interface}: All clone instances are in a class that have the same interface.
        \item \textbf{Same Class}: All clone instances are in an inheritance hierarchy that contains the same interface.
    \end{itemize}
    \item \textbf{Unrelated}: All clone instances are in classes that have the same interface.
    \begin{itemize}
        \item \textbf{No Direct Superclass}: All clone instances are in classes that have the Object class as parent.
        \item \textbf{No Indirect Superclass}: All clone instances are in a hierarchy that contains a class that has the Object class as parent.
        \item \textbf{External Superclass}: All clone instances are in classes the same external class as parent.
        \item \textbf{Indirect External Ancestor}: All clone instances are in a hierarchy that contains a class that has an external class as parent.
    \end{itemize}
\end{itemize}

Based on these relations, we determine where to place the cloned code when refactored. The code of clones that have a \textit{Common Class} relation can be refactored by placing the cloned code in this same class. The code of clones with a \textit{Common Hierarchy} relation can be placed in the intersecting class in the hierarchy (the class all clone instances have in common as an ancestor). The code of clones with a \textit{Common Interface} relation can be placed in the intersecting interface, but in the process has to become part of the classes' public contract. The code of clones that are \textit{Unrelated} can be placed in a newly created place: either a utility class, a new superclass abstraction or an interface.

\subsubsection{Contents}
% The rationale for our categories regarding clone contents.
The contents of a clone instance determine what refactoring techniques can be applied to refactor such clones. We define the following categories by which we analyze the contents of clones:
\begin{enumerate}
  \item \textbf{Full Method/Constructor/Class/Interface/Enumeration:} A clone that spans a full class, method, constructor, interface or enumeration, including its declaration.
  \item \textbf{Partial Method/Constructor:} A clone that spans (a part of) the body of a method/constructor.
  \item \textbf{Several Methods:} A clone that spans over two or more methods, either fully or partially, but does not span anything but methods.
  \item \textbf{Only Fields:} A clone that spans only global variables.
  \item \textbf{Other:} Anything that does not match with above-stated categories.
\end{enumerate}

\subsubsection{Full Method/Constructor/Class/Interface/Enumeration}
These categories denote that a full declaration, including its body, is cloned with another declaration. These categories often denote redundancy and are often easy to solve: one of both declarations is redundant and should be removed. All usages of the removed declaration should be redirected to the clone instance that was not removed. We check for clones in this category by checking that the first node in a clone instance is a declaration and the last node in the clone instance is the last node in the body of that declaration.

\subsubsection{Partial Method/Constructor}
These categories describe clone instances which are found in the body of a method or constructor. These clones can often be refactored by extracting a new method out of the cloned code. We detect such clones by checking that each node in the clone instance has a ancestor node that is a MethodDeclaration, but none of the nodes is a method declaration. We describe the procedure by with we recursively seek the parent nodes of a node in Section~\ref{sec:methodlevelcr}.

\subsubsection{Several Methods}
Several methods cloned in a single class is a strong indication of implicit dependencies between two classes. This increases the chance that these classes are missing some form of abstraction, or their abstraction is used inadequately. We detect such clones by checking that each cloned node has a parent node that is a MethodDeclaration. This is the same process as described in equation \ref{eq:samemethod}. Additionally, we check that the MethodDeclaration ancestor of the first node in the clone instance differs from the last, to be sure that the clone instance spans over at least two methods.

\subsubsection{Only Fields}
This category denotes that the clone spans over only global variables, fields that are declared outside of a method. This is to indicate data redundancy: pieces of data have an implicit dependency. In such cases, these fields may have to be encapsulated in a new object. Or, the fields should be somewhere in the inheritance structure where all objects containing the clone can access them. We check for this category by validating that all nodes in the clone instance are a VariableDeclarator. Because VariableDeclarators in methods would already fall into the ``Partial Method'' category, this includes only globally defined fields.

\subsubsection{Other}
The ``Other'' category denotes all configurations of clone contents that do not fall into above categories. Often, these are combinations of above states concepts. For instance, a combination of constructors and methods is clones, or a combination of fields and methods. Such clones indicate, like the ``Several Methods'', the requirement of performing a more architectural-level refactoring. These are often more complicated, especially when aiming to automate this process.
